The letter

### 2010-06-29 ltx/PLoS/LaTeX_Manuscript.tex

In this paper, we demonstrate for the first time how intracortical connectivity can be inferred from data, based on a variant of the WC neural field model~\cite{Wilson1973}. This work provides a fundamental link between the theoretical advances in neural field modelling and patient-specific data. The paper proceeds by first deriving the continuum neural field equations in Section~\ref{NeuralModelSection}. Then a finite dimensional neural field model is derived. The model is reduced by approximating the neural field using a set of continuous basis functions, weighted by a finite dimensional state vector. Section~\ref{SpectralAnalysisSection} establishes conditions using spatial frequency analysis for both sensor and basis function spacing and width, such that the dominant dynamics of the neural field can be represented by the reduced model. The state and parameter estimation procedure is described in Section~\ref{StateAndParameterEstimationSection}. The results for the spatial frequency analysis and parameter estimation are then presented in Section~\ref{ResultsSection}. The implications and limitations of this framework are discussed in Section~\ref{DiscussionSection} along with planned future developments.

@@

This paper has presented a method for extracting patient-specific models of synaptic dynamics from electrophysiological data. In so doing, we have highlighted important aspects of the experiment design process, demonstrating the relationship between sensor bandwidth and the complexity of the field that can be inferred. The mathematical formulation we have presented scales well with the number of sensors, and we have derived a finite-order state-space representation which allows us to perform inference efficiently.

@@

Generating patient-specific neural field models has the potential to have significant impact on several areas of neuroscientific research and clinical neurology. Specifically, the development of a patient-specific state-space model will allow for the application of techniques from engineering to be applied to neuro-dynamics. For example, state tracking could be used in brain-computer interfaces or epileptic seizure prediction, where specific regions of the state-space may indicate intent of movement or imminent seizures, respectively. In addition, a patient-specific state-space model will allow application of control engineering techniques to robustly prevent the neuro-dynamics entering pathological regions of state-space using electrical stimulation, thereby possibly preventing seizure onset or abating seizure activity.

@@

The development of an estimation framework also has theoretical implications, where the ability to estimate parameters will allow testing of hypotheses that have been generated in computational modelling studies. This has the potential to validate neural field models, and in doing so, provide new insights that may lead to a greater understanding of macro-scale neurodynamics. For example, using the proposed model-based framework, we can reconstruct a neural field from data and estimate parameters to check if they correspond to theoretically derived values where certain solutions can exist. In addition, the parameters and the state-space can be explored during cognitive tasks and states of arousal such as during sleep and anaesthesia.

@@

\subsection*{Computational Complexity}

This is the first paper to derive a continuum neural field model-based estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}. The closest competitor to our `grey-box' integro-difference equation based approach is perhaps the multivariate auto-regressive (MVAR) model, which is a `black-box' approach. Examples of MVAR models for estimating functional connectivity include Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}. A distinct advantage of the integro-difference equation based technique proposed in this current study is that the number of parameters required to model the system is considerably less than a typical MVAR model. Furthermore, by using the neural field equations as a parametric form of a model, the parameters have physical meaning and, therefore, can be clinically relevant. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in the longer range cortico-cortical fibres.

>>>>

In this paper, we demonstrate for the first time how intracortical connectivity can be inferred from data, based on a variant of the WC neural field model~\cite{Wilson1973}. This work provides a fundamental link between the theoretical advances in neural field modelling and electrophysiological data. The paper proceeds by first deriving the continuum neural field equations in Section~\ref{NeuralModelSection}. Then a finite dimensional neural field model is derived. The model is reduced by approximating the neural field using a set of continuous basis functions, weighted by a finite dimensional state vector. Section~\ref{SpectralAnalysisSection} establishes conditions using spatial frequency analysis for both sensor and basis function spacing and width, such that the dominant dynamics of the neural field can be represented by the reduced model. The state and parameter estimation procedure is described in Section~\ref{StateAndParameterEstimationSection}. The results for the spatial frequency analysis and parameter estimation are then presented in Section~\ref{ResultsSection}. The implications and limitations of this framework are discussed in Section~\ref{DiscussionSection} along with planned future developments.

@@

This paper has presented a method for creating neural field models from electrophysiological data. In so doing, we have highlighted important aspects of the experiment design process, demonstrating the relationship between sensor bandwidth and the complexity of the field that can be inferred. The mathematical formulation we have presented scales well with the number of sensors, and we have derived a finite-order state-space representation which allows us to perform inference efficiently.

@@

Generating data-driven neural field models has the potential to have significant impact on several areas of neuroscientific research and clinical neurology. Specifically, the development of a data-driven state-space model will allow for the application of techniques from engineering to be applied to neuro-dynamics. For example, state tracking could be used in brain-computer interfaces or epileptic seizure prediction, where specific regions of the state-space may indicate intent of movement or imminent seizures, respectively. In addition, a patient-specific state-space model will allow application of control engineering techniques to robustly prevent the neuro-dynamics entering pathological regions of state-space using electrical stimulation, thereby possibly preventing seizure onset or abating seizure activity.

@@

The development of an estimation framework also has theoretical implications, where the ability to estimate parameters will allow testing of hypotheses that have been generated in computational modelling studies. This has the potential to validate neural field models, and in doing so, provide new insights that may lead to a greater understanding of macro-scale neurodynamics. For example, using the proposed model-based framework, we can reconstruct a neural field from data and estimate parameters to check if they correspond to theoretically derived values where certain solutions can exist. In addition, the parameters and the state-space can be explored during cognitive tasks and states of arousal such as sleep and anaesthesia.

@@

\subsection*{Computational Complexity}
This is the first paper to derive a continuum neural field model-based estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}. The closest competitor to our `grey-box' integro-difference equation based approach is perhaps the multivariate auto-regressive (MVAR) model, which is a `black-box' approach. Examples of MVAR models for estimating functional connectivity include Granger causality~\cite{Hesse2003}, the direct transfer function~\cite{Kaminski1991} and partial directed coherence~\cite{Sameshima1999}. A distinct advantage of the integro-difference equation based technique proposed in this current study is that the number of parameters required to model the system is considerably less than a typical MVAR model. Furthermore, by using the neural field equations as a parametric form of a model, the parameters have physical meaning and, therefore, can be clinically relevant. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in the longer range cortico-cortical fibres.
